---
title: "Data Tidying"
author: "young-robertson"
date: "2/16/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in library packages at the beginning
Message, warning = FALSE suppresses warnings in rendered document

to install packages, do this:
install.packages("dplyr")
But it's already been done on this server

```{r, message = FALSE, warning = FALSE}

library(dplyr)
library(tidyr)
library(readr)

```

Sometimes have to identify which function you're using from which package -- 
package_name::function_name

like this:
stats::filter()

https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1

## Clean Data

Piping to string together functions to apply to the data
 %>% : Ctrl + Shift + M (Windows)

```{r}

catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

#or can do this:

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"))

head(catch_original)
str(catch_original)

```

```{r}
catch_data <- catch_original %>% 
  #select(Region , Year, Chinook, Sockeye, Coho, Pink, Chum)  #selects the columns we want to keep
  select(-All, -notesRegCode)  #also selects the columns we want to keep, so we're not keeping All and notesRegCode

head(catch_data)
glimpse(catch_data)
summary(catch_data)
#summary is showing us Chinook is a Chr data types

```

-change chinook from chr to num

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1" , Chinook)) %>%  #changes chr I to chr 1 -- did this after running line 81 and 83
  mutate(Chinook = as.numeric(Chinook))  #changes Chinook from chr to num
  
head(catch_clean)
str(catch_clean)
```

### finding spurious characters in Chinook column
```{r}
#which returns row numbers or indices for which a statement is true
(i = which(is.na(catch_clean$Chinook)))  #only row 401 is the problem , is.na identifies where there are NA's with TRUE or FALSE

catch_original[i,] #shows us the offending row, shows an "I" instead of a 1

```

### reshaping data
-move from wide to long format

```{r}
catch_long = catch_clean %>% 
  pivot_longer(cols = -c(Region, Year) , names_to = "Species" , values_to = "Catch") %>% 
  rename(catch_thousands = Catch) %>%  #renames Catch to catch_thousands
  mutate(catch = catch_thousands *1000) %>% 
  select(-catch_thousands)
  
#not going to pivot over region and year (see the "-" before "c"), column we're going to create (takes the name of the columns and puts them in Species column, takes the values and puts them in the catch column)
#put quotes around the new variable or name, don't need to use quotes if we're calling a variable that already exists

head(catch_long)
```


```{r}
catch_wide = catch_long %>% 
  pivot_wider(names_from = Region, values_from = catch)  #create columns from region and populate it with catch data

head(catch_wide)
```


## Split, combine, apply

- calculate mean salmon catch by region

```{r}
mean_region = catch_long %>% 
  filter(Species == "Sockeye") %>% #just shows us sockeye catch mean data
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(catch) , n = n() , .groups = "drop")  #n function counts number of rows in each observations

head(mean_region)

```

### Joining

```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```


```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))  #use "by" argument because there isn't a common column between the two datasets

head(catch_joined)
nrow(catch_joined) == nrow(catch_long)  #True
```
