---
title: "workshop_Rmarkdown"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

## Header 2

### Header 3

Write some sample text here. **Bold text**

Use the Knit button to show what the rendered version looks like

Create lists using the single \* or a - and get bullets, need to include a blank line after this line

-   apple
-   banana

# First Challenge

Write some *italic* text

#### **Second Challenge**

Numbered list

1.  apple
2.  banana
3.  kiwi

This will order it for you (as 1, 2, 3)

1.  apple
2.  banana
3.  kiwi

Can use "markdown quick reference" from the Help menu to get some quick tips or cheat sheet

Can use the compass symbol on the top right to have it format as you go (visual markdown editor). Good for adding citations.

To insert code chunk
cmd + option + i (on mac)

cntrl + alt + i (on pc)

```{r}

x = 3+2

(y = sum(4, 5, 6, 7))

```

## Read in data

../ goes up a folder, hit "tab" and will show you the options from which to select

```{r}

bg_chem = read.csv("../data/BGchem2008data.csv")

```


## Analysis

## Plots

echo = TRUE means that the R code will show up in the rendered document. The default from the beginning is TRUE

Help -> cheat sheets -> R markdown reference guide or cheat sheet

```{r , echo=TRUE}

plot(bg_chem$CTD_Depth , bg_chem$CTD_Temperature)

```
# Discussion
This is my first introduction to Git and I think it would be a great tool to use with my new grad student. We'll be working on plant phys models in OpenBUGS, so I will have to sort out how to use Git with this program (.odc files). This tutorial today has really demystified Git for me but I can see that I'll have to refer back to the instructions a few times to catch on.

