---
title: "data-visualization"
author: "young-robertson"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warnings = FALSE}
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
  library(scales) # install.packages("scales")
```

#do the following in case it can't retrieve data from the url, then can grab it from elsewhere

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv"),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"))
        return(esc)
    }
)

head(esc)

```

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)
```

## ggplot

```{r}
ggplot(annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col(position = "dodge")

#dodge breaks up the single columns/fill values into bars that are next to each other
#stack stacks them on top of each other (default is stacking)

```


```{r}
kodiak_esc <- annual_esc %>% 
  filter(SASAP.Region == "Kodiak")
```

```{r}
#can create your own theme for the plot aesthetics
my_theme = theme_bw() +
  theme(legend.position = "bottom")

ggplot(kodiak_esc, aes(x=Year, y=escapement, color = Species)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = comma) +  #ylabel is no longer in sci notation
  scale_x_continuous(n.breaks = 10) +  #adds extra major ticks to x axis
  scale_color_brewer(palette="Dark2") +
  ylab("Escapement (num fish)") +
  ggtitle("Kodiak Escapement") +
  my_theme
  
  #theme(legend.position = "bottom")   #can change every part of the plot

```

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()  #drop any row with an NA
```

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```
Might need to generate figures as html to have figures pop up for each site
might be addPopups , put html in the addPopups() , can create a column in data with html link, but could make website loading slow, so could just put links to graphs in the popup rather than the graphs themselves
